---
title: "BELLABEAT CASE STUDY"
author: "Nicholas Newdorf"
date: "2022-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
install.packages("tidyverse")
library(tidyverse)
setwd("./Data")

daily_activity <- read_csv("dailyActivity_merged.csv")
daily_activity_filtered <- daily_activity %>%
  filter(!(TotalSteps == 0 & SedentaryMinutes == 1440))

sleep_day <- read_csv("sleepDay_merged.csv")
weight_log_info <- read_csv("weightLogInfo_merged.csv")
heartrate_seconds <- read_csv("heartrate_seconds_merged.csv")
minute_mets_narrow <- read_csv("minuteMETsNarrow_merged.csv")
```
## THE TASK
Determine suggestions for Bellabeat marketing by analyzing current trends in other smart trackers.

* What current Bellabeat offerings match user trends?
* Are there deficiencies in the current line up?

## THE DATA
A data set was sourced over the course of a month from 33 Fitbit users

### Data Limitations
* Only 33 participants in the study
* Bellabeat caters to a female audience, the gender of Fitbit users is unknown
* There is variance in the Fitbit types leading to subsets of participants for some metrics

### Data Cleaning Process
72 rows with 0 steps and 1440 sedentary minutes were removed under the assumption that this indicated days the device was not worn

## THE ANALYSIS

### Time Spent Measuring Throughout The Day
Not all devices are used for a full 24 hour day.
Some participants never hit a full day while others did so consistently.
Participants seem to congregate at either one extreme or another with few in between.

```{r, echo=FALSE}
num_full_days <- daily_activity_filtered %>% 
  rowwise() %>%
  mutate(TotalMinute = sum(c(VeryActiveMinutes,FairlyActiveMinutes,LightlyActiveMinutes,SedentaryMinutes))) %>% 
  group_by(Id) %>% summarise(Count = sum(TotalMinute == 1440))

ggplot(num_full_days, aes(x=Count)) + 
  geom_bar() + 
  labs(title="Number of Full 24 Hour Days Recorded by Users", x="Number of Full Days", y="Number of Users")
```

### Types of Activities
Activities are measured in four groups: Sedentary, Lightly Active, Fairly Active, and Very Active.
Sedentary activity is by far the most measured activity.

```{r, echo=FALSE}
longer <- daily_activity_filtered %>%
  pivot_longer(VeryActiveMinutes:SedentaryMinutes, names_to = "ActivityType", values_to = "Minutes")
longer_summed <- longer %>%
  group_by(ActivityType) %>%
  summarize(Minutes = sum(Minutes))
longer_summed <- longer_summed %>% 
  mutate(Percentage = Minutes/sum(Minutes))
longer_summed <- longer_summed %>%
  mutate(PercentageLabel = scales::percent(Percentage,1))
longer_summed$ActivityType <- factor(longer_summed$ActivityType, levels=c("SedentaryMinutes", "LightlyActiveMinutes", "FairlyActiveMinutes", "VeryActiveMinutes"))


ggplot(longer_summed, aes(x="", y=Percentage, fill=ActivityType)) + 
  geom_bar(stat="identity", width=1) + 
  coord_polar("y",start=0) + 
  theme_void() + 
  scale_fill_grey() + 
  geom_text(aes(label = ifelse(Percentage >= 0.1, PercentageLabel,"")),
            position = position_stack(vjust = 0.5),
            color="white") +
  labs(title="Activity Percentages")
```

### Types of Metrics
There are multiple metrics that can be tracked by FitBits. Some are more tracked than others.

```{r, echo=FALSE}
NumUsers <- c(
n_distinct(daily_activity_filtered$Id),
n_distinct(daily_activity_filtered$Id),
n_distinct(daily_activity_filtered$Id),
n_distinct(sleep_day$Id),
n_distinct(weight_log_info$Id),
n_distinct(heartrate_seconds$Id),
n_distinct(minute_mets_narrow$Id))

Metric <- c(
  "Steps",
  "Intensities",
  "Calories",
  "Sleep",
  "Weight",
  "HeartRate",
  "MET"
)

user_of_metrics <- data.frame(Metric, NumUsers)

ggplot(user_of_metrics, aes(x=reorder(Metric, NumUsers), y=NumUsers)) +
  geom_col() +
  labs(title = "Number Of Distinct Users For A Given Metric", x="Metric", y="Number of Users", caption = "Note: Weight is tracked either through manual entry or a connected scale")
```

## CONCLUSION

### Key Findings
* The uses of current smart trackers vary:
  * Most of the time users are sedentary with small bursts of higher intensity activity.
  * Sleep tracking and activity tracking are widely used while fewer people track heart-rate and weight.
* Metrics that require manual input (like weight) appear to be less widely used than those automatically tracked.

### Recomendation
Out of Bellabeat's current lineup, the Leaf most closely resembles the data tracked in this data set. The Ivy includes extra features that some might use such as heart-rate tracking, but that accounts for less than half the participants of the study. For this purpose, it is my recommendation that we focus marketing on the Leaf.

### Future Analysis Opportunities
This study focused on a small set of Fitbit users and had many data limitations. Finding bigger data sets that more closely resemble Bellabeat's clientele would provide a better picture of the smart tracker landscape.

## APENDIX
Dataset taken from:
https://www.kaggle.com/datasets/arashnic/fitbit/

Data dictionary from:
https://www.fitabase.com/media/1930/fitabasedatadictionary102320.pdf

Differences between Leaf and Ivy:
https://support.bellabeat.com/hc/en-us/articles/360020587680-What-is-the-difference-between-Ivy-and-the-Leaf-